---
title: 'McKinney Fire metabolism: Time series model + breakpoint model'
author: "Laurel Genzoli"
date: "2026-02-23"
format:
  html:
    self-contained: true
    fig-align: center
    theme: united
    highlight: tango
    number-sections: true
  pdf:
    pdf-engine: xelatex
    number-sections: true
    toc: true
    fig-align: center
    keep-tex: true
editor_options: 
  markdown: 
    wrap: 72
---

```{=html}
<style type="text/css">
body, td {font-size: 12px;}
code.r{font-size: 8px;}
pre {font-size: 10px}
</style>
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = F, message = F)
knitr::opts_knit$set(root.dir = '/Users/laurelgenzoli/Dropbox/2024_Projects/Mckinney_Fire_Metab/Mck-Fire-Metab-GH')
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})

```

# Script description

This script contains the time series model testing the effects of the McKinney Fire on stream metabolism, turbidity, light, and discharge.  


### Packages
```{r, echo = T, message = F}
setwd('/Users/laurelgenzoli/Dropbox/2024_Projects/Mckinney_Fire_Metab/Mck-Fire-Metab-GH')


library(lubridate)
library(dataRetrieval)
library(viridis)
library(cowplot)
library(ggpubr)
library(streamMetabolizer)
library(hms)
library(plotly)
library(scales)
library(data.table)
library(zoo)
library(nlme)
library(tidyverse)
library(brms)
library(tidybayes)
library(segmented)

theme_set(theme_classic())
```

### Load data
```{r, echo = T, message = F}
met.mle.final <- read_csv("Data/met_mle_final.csv") %>%
 mutate(year = as.factor(year),
         fire = ifelse(date > "2022-08-02", "after", "before"))%>%
  mutate(GPP = ifelse(QA.code == 1, NA, GPP))%>%
  mutate(GPP.upper = ifelse(QA.code == 1, NA, GPP.upper))%>%
  mutate(GPP.lower = ifelse(QA.code == 1, NA, GPP.lower))%>%
  mutate(ER = ifelse(QA.code %in% c(1,2), NA, ER))%>%
  mutate(ER.upper = ifelse(QA.code %in% c(1,2), NA, ER.upper))%>%
  mutate(ER.lower = ifelse(QA.code %in% c(1,2), NA, ER.lower))%>%
  mutate(NEP = ifelse(is.na(ER), NA, NEP))%>%
  dplyr::select(-1)

names(met.mle.final)
```

# Bayes version of interrupted time series

This is a more robust test of change following the fire than a simple comparison of means because it incorporates the autocorrelation and seasonality of the time series. The intercept change at the event is obvious to our eyes, but it is nice to have a statistical test behind it, and a better estimate of the magnitude of change. 


## Prep data set for model

```{r, warning=F, size = 'small', echo=FALSE, fig.width=8, fig.height=5}

ts_data<- met.mle.final %>%
  mutate(t = as.numeric(date - min(date)),  # time index
    event = ifelse(fire == "before", 0, 1),  
    time_after_event = ifelse(event == 1, t - min(t[event==1]), 0), #slope change after event
    sin1 = sin(2*pi*jday/365.25),
    cos1 = cos(2*pi*jday/365.25))%>% #harmonic ex (pair captures annual cycle)
    dplyr::select(t, event, time_after_event, mon, sin1, cos1, turb_median, SV.cms, GHI.day, GPP, ER, NEP)
```   


## Fit the model with AR(1): Formula defined in code below

```{r, results = 'hide', warning=F, size = 'small', echo=FALSE, fig.width=8, fig.height=5}
fit.q <- brm(formula = bf(SV.cms ~ t + event  + sin1 + cos1),
  data = ts_data,
  family = gaussian(),
  autocor = cor_ar(~ t, p = 1),  
  chains = 4, cores = 4, iter = 1000,
  control = list(adapt_delta = 0.95))

fit.l <- brm(formula = bf(GHI.day ~ t + event + sin1 + cos1),
  data = ts_data,
  family = gaussian(),
  autocor = cor_ar(~ t, p = 1),  
  chains = 4, cores = 4, iter = 1000,
  control = list(adapt_delta = 0.95))

fit.turb <- brm(formula = bf(turb_median ~ t + event +  sin1 + cos1),
  data = ts_data,
  family = gaussian(),
  autocor = cor_ar(~ t, p = 1),  
  chains = 4, cores = 4, iter = 1000,
  control = list(adapt_delta = 0.95))

fit.gpp <- brm(formula = bf(GPP ~ t + event + sin1 + cos1),
  data = ts_data,
  family = gaussian(),
  autocor = cor_ar(~ t, p = 1), 
  chains = 4, cores = 4, iter = 1000,
  control = list(adapt_delta = 0.95))

fit.er <- brm(formula = bf(ER ~ t + event  + sin1 + cos1),
  data = ts_data,
  family = gaussian(),
  autocor = cor_ar(~ t, p = 1), 
  chains = 4, cores = 4, iter = 1000,
  control = list(adapt_delta = 0.95))

fit.nep <- brm(formula = bf(NEP ~ t + event  + sin1 + cos1),
  data = ts_data,
  family = gaussian(),
  autocor = cor_ar(~ t, p = 1),  
  chains = 4, cores = 4, iter = 1000,
  control = list(adapt_delta = 0.95))
```


## Inspect posterior summaries


```{r, warning=F, size = 'small', echo=FALSE, fig.width=8, fig.height=5}

summary(fit.q)
summary(fit.l)
summary(fit.turb)
summary(fit.gpp)
summary(fit.er)
summary(fit.nep)


psq <- posterior_summary(fit.q)
psl <-posterior_summary(fit.l)
pst <-posterior_summary(fit.turb)
psg <-posterior_summary(fit.gpp)
pse <-posterior_summary(fit.er)
psn <-posterior_summary(fit.nep)

fits <- list(psq, psl, pst, psg, pse, psn)



```



## Extract means pre/post fire for text and plots
   
    - These are means while holding season and trend stable
    - Time trend is average, but sin and cos are arbitrarily set to 0
    - This is data used in figure 2, right panels

```{r, warning=F, size = 'small', echo=FALSE, fig.width=8, fig.height=5}

df.l <- tibble(event = c(0,1),t = mean(ts_data$t), sin1 = 0, cos1 = 0) %>%
  rowwise() %>%
  mutate(pred = list(posterior_epred(fit.l, newdata = cur_data()))) %>%
  mutate(mean = mean(unlist(pred)),
         lower = quantile(unlist(pred), 0.025),
         upper = quantile(unlist(pred), 0.975),
         group = ifelse(event==0,"Before","After")) %>%
  dplyr::select(group, mean, lower, upper)

df.q <- 
  tibble(event = c(0,1),t = mean(ts_data$t),sin1 = 0, cos1 = 0) %>%
  rowwise() %>%
  mutate(pred = list(posterior_epred(fit.q, newdata = cur_data()))) %>%
  mutate(mean = mean(unlist(pred)),
         lower = quantile(unlist(pred), 0.025),
         upper = quantile(unlist(pred), 0.975),
         group = ifelse(event==0,"Before","After")) %>%
  dplyr::select(group, mean, lower, upper)

df.t <- 
  tibble(event = c(0,1),t = mean(ts_data$t),sin1 = 0, cos1 = 0) %>%
  rowwise() %>%
  mutate(pred = list(posterior_epred(fit.turb, newdata = cur_data()))) %>%
  mutate(mean = mean(unlist(pred)),
         lower = quantile(unlist(pred), 0.025),
         upper = quantile(unlist(pred), 0.975),
         group = ifelse(event==0,"Before","After")) %>%
  dplyr::select(group, mean, lower, upper)

df.g <- 
  tibble(event = c(0,1),t = mean(ts_data$t),sin1 = 0, cos1 = 0) %>%
  rowwise() %>%
  mutate(pred = list(posterior_epred(fit.gpp, newdata = cur_data()))) %>%
  mutate(mean = mean(unlist(pred)),
         lower = quantile(unlist(pred), 0.025),
         upper = quantile(unlist(pred), 0.975),
         group = ifelse(event==0,"Before","After")) %>%
  dplyr::select(group, mean, lower, upper)

df.e <- 
  tibble(event = c(0,1),t = mean(ts_data$t),sin1 = 0, cos1 = 0) %>%
  rowwise() %>%
  mutate(pred = list(posterior_epred(fit.er, newdata = cur_data()))) %>%
  mutate(mean = mean(unlist(pred)),
         lower = quantile(unlist(pred), 0.025),
         upper = quantile(unlist(pred), 0.975),
         group = ifelse(event==0,"Before","After")) %>%
  dplyr::select(group, mean, lower, upper)

df.n <- 
  tibble(event = c(0,1),t = mean(ts_data$t),sin1 = 0, cos1 = 0) %>%
  rowwise() %>%
  mutate(pred = list(posterior_epred(fit.nep, newdata = cur_data()))) %>%
  mutate(mean = mean(unlist(pred)),
         lower = quantile(unlist(pred), 0.025),
         upper = quantile(unlist(pred), 0.975),
         group = ifelse(event==0,"Before","After")) %>%
  dplyr::select(group, mean, lower, upper)

```





## Means to report in text (make a table)
    
    - Essentially, prediction of each day, added up across time series
    - This incorporates seasonality and trend into means estimates

```{r, warning=F, size = 'small', echo=FALSE, fig.width=8, fig.height=5}
before_df <- ts_data %>% filter(event == 0)
after_df  <- ts_data %>% filter(event == 1)

pred_b <- posterior_epred(fit.q, newdata = before_df)
pred_a <- posterior_epred(fit.q, newdata = after_df)
m.q <- tibble(group = c("Before","After"),
  mean  = c(mean(pred_b), mean(pred_a)),
  lower = c(quantile(pred_b,0.025), quantile(pred_a,0.025)),
  upper = c(quantile(pred_b,0.975), quantile(pred_a,0.975)))%>%
  mutate(param = "flow")


pred_b <- posterior_epred(fit.l, newdata = before_df)
pred_a <- posterior_epred(fit.l, newdata = after_df)
m.l <- tibble(group = c("Before","After"),
  mean  = c(mean(pred_b), mean(pred_a)),
  lower = c(quantile(pred_b,0.025), quantile(pred_a,0.025)),
  upper = c(quantile(pred_b,0.975), quantile(pred_a,0.975)))%>%
  mutate(param = "light")

pred_b <- posterior_epred(fit.turb, newdata = before_df)
pred_a <- posterior_epred(fit.turb, newdata = after_df)
m.t <- tibble(group = c("Before","After"),
  mean  = c(mean(pred_b), mean(pred_a)),
  lower = c(quantile(pred_b,0.025), quantile(pred_a,0.025)),
  upper = c(quantile(pred_b,0.975), quantile(pred_a,0.975)))%>%
  mutate(param = "turb")

pred_b <- posterior_epred(fit.gpp, newdata = before_df)
pred_a <- posterior_epred(fit.gpp, newdata = after_df)
m.g <- tibble(group = c("Before","After"),
  mean  = c(mean(pred_b), mean(pred_a)),
  lower = c(quantile(pred_b,0.025), quantile(pred_a,0.025)),
  upper = c(quantile(pred_b,0.975), quantile(pred_a,0.975)))%>%
  mutate(param = "gpp")

pred_b <- posterior_epred(fit.er, newdata = before_df)
pred_a <- posterior_epred(fit.er, newdata = after_df)
m.e <- tibble(group = c("Before","After"),
  mean  = c(mean(pred_b), mean(pred_a)),
  lower = c(quantile(pred_b,0.025), quantile(pred_a,0.025)),
  upper = c(quantile(pred_b,0.975), quantile(pred_a,0.975)))%>%
  mutate(param = "er")

pred_b <- posterior_epred(fit.nep, newdata = before_df)
pred_a <- posterior_epred(fit.nep, newdata = after_df)
m.n <- tibble(group = c("Before","After"),
  mean  = c(mean(pred_b), mean(pred_a)),
  lower = c(quantile(pred_b,0.025), quantile(pred_a,0.025)),
  upper = c(quantile(pred_b,0.975), quantile(pred_a,0.975)))%>%
  mutate(param = "nep")

bind_rows(m.l, m.q, m.t, m.g, m.e, m.n)
```



## Plot fitted values vs. observed


```{r, warning=F, size = 'small', echo=FALSE, fig.width=8, fig.height=10}

ts_data$fitted.turb  <- fitted(fit.turb, newdata = ts_data, re_formula = NA)[, "Estimate"]
ts_data$fitted.gpp   <- fitted(fit.gpp, newdata = ts_data, re_formula = NA)[, "Estimate"]
ts_data$fitted.er <- fitted(fit.er, newdata = ts_data, re_formula = NA)[, "Estimate"]
ts_data$fitted.nep   <- fitted(fit.nep, newdata = ts_data, re_formula = NA)[, "Estimate"]


t <- ggplot(ts_data, aes(x = t)) +
  geom_line(aes(y = turb_median), color = "grey60") +
  geom_line(aes(y = fitted.turb), color = "red", lwd = 0.3) +
  geom_vline(xintercept = 1540, linetype = "dashed", color = "orange", lwd = 1.2) +
  labs(y = "Turbidity",
    x = "Time",
    title = "Interrupted Time Series with Seasonality & AR(1)")

g <- ggplot(ts_data, aes(x = t)) +
  geom_line(aes(y = GPP), color = "grey60") +
  geom_line(aes(y = fitted.gpp), color = "red", lwd = 0.3) +
  geom_vline(xintercept = 1540, linetype = "dashed", color = "orange", lwd = 1.2) +
  labs(y = "GPP",
    x = "Time",
    title = "Interrupted Time Series with Seasonality & AR(1)")

r <- ggplot(ts_data, aes(x = t)) +
  geom_line(aes(y = ER), color = "grey60") +
  geom_line(aes(y = fitted.er), color = "red", lwd = 0.3) +
  geom_vline(xintercept = 1540, linetype = "dashed", color = "orange", lwd = 1.2) +
  labs(y = "ER",
    x = "Time")

n <- ggplot(ts_data, aes(x = t)) +
  geom_line(aes(y = NEP), color = "grey60") +
  geom_line(aes(y = fitted.nep), color = "red", lwd = 0.3) +
  geom_vline(xintercept = 1540, linetype = "dashed", color = "orange", lwd = 1.2) +
  labs(y = "NEP",
    x = "Time")

plot_grid(t, g, r, n, ncol = 1)

```





# A GLM with a turbidity breakpoint for dissolved oxygen response

- phi = 0, including autocorrelation did not help so using GLM
- This still includes both the gls and glm model below
- The final model and the figure are in the figure script


```{r, warning=F, size = 'small', echo=FALSE, fig.width=8, fig.height=4}

gls.data <- met.mle.final%>% 
  filter(min.DO != -Inf)%>%
  filter(turb_median < 50)%>%
  mutate(min.DO = min.DO*100, t = log(turb_median), 
         q = log(SV.cms), l = log(GHI.day), g = log(GPP))

## gls: no autocorrelation here so can use GLM
gls_minDO <- gls(min.DO ~ t + q + l,
                 data = gls.data,
                 na.action = na.omit,
                 correlation = corAR1(form = ~ 1 | date)) 
summary(gls_minDO)

## quick figure
gls.data %>%
  mutate(mon = month(date))%>%
  ggplot(aes(x  = t, y = min.DO, col = l))+ #fire, year, etc
           geom_point()

## GLM (no autocorrelation)
bp.mod <- glm(min.DO ~ t + q + l,
              data = gls.data,
              na.action = na.omit)

seg_mod <- segmented(bp.mod, seg.Z = ~t, 
                     psi = list("t" = 10))
summary(seg_mod)
plot(seg_mod)
resid_var <- var(residuals(gls_minDO))
total_var <- var(gls.data$min.DO)
R2 <- 1 - resid_var / total_var
R2

```




