---
title: 'McKinney Fire metabolism: Publication figures'
author: "Laurel Genzoli"
date: "2026-02-23"
format:
  html:
    self-contained: true
    fig-align: center
    theme: united
    highlight: tango
    number-sections: true
  pdf:
    pdf-engine: xelatex
    number-sections: true
    toc: true
    fig-align: center
    keep-tex: true
editor_options: 
  markdown: 
    wrap: 72
---

```{=html}
<style type="text/css">
body, td {font-size: 12px;}
code.r{font-size: 8px;}
pre {font-size: 10px}
</style>
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = F, message = F)
knitr::opts_knit$set(root.dir = '/Users/laurelgenzoli/Dropbox/2024_Projects/Mckinney_Fire_Metab/Mck-Fire-Metab-GH')
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})

```



# Script Description

This script make the publication plots for the first draft of the McKinney Fire Metabolism paper, which is the report. Some plots were post processed to lay out (figure 1).

## Load packages and data

### Packages
```{r, echo = T, message = F}
setwd('/Users/laurelgenzoli/Dropbox/2024_Projects/Mckinney_Fire_Metab/Mck-Fire-Metab-GH')


library(lubridate)
library(dataRetrieval)
library(viridis)
library(cowplot)
library(ggpubr)
library(streamMetabolizer)
library(hms)
library(plotly)
library(scales)
library(data.table)
library(zoo)
library(nlme)
library(tidyverse)
library(magick)
library(scales)
library(ggbreak)
library(brms)
library(segmented)
library(ggnewscale)


theme_set(theme_classic())
```


## Data

Data including both daily outputs and sub-daily/high frequency data.

```{r, echo = T, message = F}

## Final metabolism data, with QA points turned to NA
met.mle.final <- read_csv("Data/met_mle_final.csv") %>%
 mutate(year = as.factor(year),
         fire = factor(ifelse(date > "2022-08-02", "After", "Before"),
                       levels = c("Before", "After"))) %>%
  mutate(GPP = ifelse(QA.code == 1, NA, GPP))%>%
  mutate(GPP.upper = ifelse(QA.code == 1, NA, GPP.upper))%>%
  mutate(GPP.lower = ifelse(QA.code == 1, NA, GPP.lower))%>%
  mutate(ER = ifelse(QA.code %in% c(1,2), NA, ER))%>%
  mutate(ER.upper = ifelse(QA.code %in% c(1,2), NA, ER.upper))%>%
  mutate(ER.lower = ifelse(QA.code %in% c(1,2), NA, ER.lower))%>%
  mutate(NEP = ifelse(is.na(ER), NA, NEP))
  

## Final metabolism data day of pulse
met.mle.final.FIRE <- read_csv("Data/met_mle_final.csv")
  

## Data for sub-daily response plots during Fire
sub.day.1 <- read_csv("Data/sub-daily-FIRE.csv")%>%
  mutate(datetime = force_tz(datetime, tzone = "Etc/GMT+7"))

sub.day.do <- read_csv("Data/SV_clean-23072025.csv")%>%
  mutate(datetime = force_tz(solar.time, tzone = "Etc/GMT+7"))



## model fit data used in SI turbidity spike plot
mod.fits.bin <- read_csv("Data/Metab_Output/Binned/SV_fits_BIN.csv")%>%
 mutate(week = week(solar.time))%>%
 mutate(jday = yday(solar.time),
        year = year(solar.time),
        mon = month (solar.time),
        DO.def = DO.sat - DO.obs)%>%
   mutate(solar.time = force_tz(solar.time, tz = "Etc/GMT+7"))


## Sub daily turbidity: used in spike plots and in stats
## This data is not included in the public GIT Release
## Contact Karuk Tribe DNR for this data

hour.turb <- read_csv("Data/SV_2018-2024_AllParams.csv", 
                      skip = 16, 
                      col_names = c("dtime", "temp.water", "sc", "domg", "ph", "turb")) %>%
  mutate(dtime = ymd_hms(dtime)) %>%                 # convert to POSIXct
  separate(dtime, into = c("date", "time"), sep = " ", remove = F) %>%
  mutate(date = as.Date(date),
        datetime = force_tz(dtime, tz = "Etc/GMT+7"),
         solar.time = datetime) %>%                  # use the datetime column
  filter(date >= as.Date("2022-07-30")) %>%
  dplyr::select(solar.time, turb)


hour.turb.all <- read_csv("Data/SV_2018-2024_AllParams.csv", 
                      skip = 16, 
                      col_names = c("dtime", "temp.water", "sc", "domg", "ph", "turb")) %>%
  mutate(dtime = ymd_hms(dtime)) %>%                 # convert to POSIXct
  separate(dtime, into = c("date", "time"), sep = " ", remove = F) %>%
  mutate(date = as.Date(date),
        datetime = force_tz(dtime, tz = "Etc/GMT+7"),
         solar.time = datetime) %>%                  # use the datetime column
  dplyr::select(solar.time, turb)

```


# Figures for main manuscript




## Figure 1: During the storm panel

  - Instantaneous light, flow, turb, DO for 6 days
  - To plot below with the map and photos of event
  - 1 sub-panel for each parameter

```{r, warning=F, size = 'small', echo=FALSE, fig.width=4, fig.height=5.5}
sub.day <- sub.day.1 %>% slice(-c(1:6))

l <- sub.day %>% 
  filter(!is.na(GHI))%>%
  ggplot(aes(x = datetime, y = GHI_interp))+
  annotate( "rect",xmin = as.POSIXct("2022-08-02 17:30", tz = "Etc/GMT+8"), 
    xmax = as.POSIXct("2022-08-02 20:30" , tz = "Etc/GMT+8"),
    ymin = -Inf, ymax = Inf,
    fill = "grey70", alpha = 0.3) +
  geom_line(col = "gold2", lwd = 1.2, alpha = 0.8)+
  geom_point(col = "gold2", size = 0.6)+
  ylab(expression(Surface~light~(W~m^-2)))+
  xlab(NULL) +                                  
  scale_x_datetime(date_breaks = "1 day",  date_labels = "",  expand = c(0, 0)) +   scale_y_continuous(limits = c(0, 1140))+
 theme( axis.title.y = element_text(size = 9),
    axis.text.x = element_blank(),  # remove all x-axis labels
    axis.ticks.x = element_line(),  # keep ticks
    plot.margin = margin(t = 0, r = 0, b = -20, l = 0))+
  annotate("text",
         x = as.POSIXct("2022-08-02 17:00", tz = "Etc/GMT+8"),  
         y = 1050,  
         label = "Intense rainfall on\n active wildfire",
         hjust = 1.05,  
         color = "grey20", size = 3)

q <- sub.day %>% 
  ggplot(aes(x = datetime, y = SV.cfs*0.0283))+
  geom_line(col = "cyan4", lwd = 1.2, alpha = 0.8)+
  ylab(expression(Discharge~(m^3~s^-1)))+
  xlab(NULL) +                                  
  scale_x_datetime(date_breaks = "1 day",  date_labels = "",  expand = c(0, 0)) +     theme( axis.title.y = element_text(size = 9),
    axis.text.x = element_blank(),  # remove all x-axis labels
    axis.ticks.x = element_line(),  # keep ticks
    plot.margin = margin(t = 0, r = 0, b = -2, l = 0))+
   annotate("text",
         x = as.POSIXct("2022-08-03 02:00", tz = "Etc/GMT+8"),  
         y = 52,  
         label = "Flow pulse arrived to\n Seiad 9 hr after\n upriver rains",
         hjust = 1.05,  
         color = "grey20", size = 3)

t <- sub.day %>% 
  ggplot(aes(x = datetime, y = turb))+
  geom_line(col = "chocolate4", lwd = 1.2, alpha = 0.8)+
  ylab("Turbidity (NTU)")+
  xlab(NULL) +  
 # scale_y_log10()+
  scale_x_datetime(date_breaks = "1 day",  date_labels = "",  expand = c(0, 0)) + theme( axis.title.y = element_text(size = 9),
    axis.text.x = element_blank(),  # remove all x-axis labels
    axis.ticks.x = element_line(),  # keep ticks
    plot.margin = margin(t = 0, r = 2, b = -2, l = 2))+
   annotate("text",
         x = as.POSIXct("2022-08-03 11:00", tz = "Etc/GMT+8"),  
         y = 530,  
         label = "First large turbidity peak\n arrived 18 hr after\n upriver rains",
         hjust = 1.12,  
         color = "grey20", size = 3)


do <- sub.day %>%
  ggplot(aes(x = datetime, y = domg)) +
  geom_line(col = "skyblue", lwd = 1.2, alpha = 0.8) +
  ylab(expression(Dissolved~O[2]~(mg~L^-1))) +
  xlab(NULL) +
  scale_x_datetime(date_breaks = "1 day", expand = c(0,0), date_labels = "%d-%b") +
  theme(axis.title.y = element_text(size=9),
        axis.ticks.x = element_line(),
        plot.margin = margin(t=0,r=0,b=-20,l=0)) +
  annotate("text",
           x = as.POSIXct("2022-08-03 16:00", tz="Etc/GMT+8"),
           y = 1.5,
           label = expression(atop(Dissolved~O[2]~plummeted, "in-step with"~peak~turbidity)),
           hjust = 1.1,
           color = "grey20", size = 3)

event <- plot_grid(l, q, t, do, nrow = 4, align = "hv", labels = c("b", "c", "d", "e"),
                   label_x = 0.95, label_y = 1.05 )
#event <- event + theme(plot.margin=margin(10,-20,0,0))

#pdf("Figures/F1_Map/bcde.pdf", height = 5.5, width = 4)
event
#dev.off()
```




## Figure 2: Full time series showing change after fire

### Left panels, time series of light, flow, turb and metabolism

```{r, warning=F, size = 'small', echo=FALSE, fig.width=5, fig.height=8}

unique(met.mle.final$fire)

l <- ggplot(met.mle.final, aes(x = date, y = GHI.day, alpha = fire))+
  geom_vline(xintercept = as.Date("2022-08-03"), lty = 2, col = "black")+
  geom_vline(xintercept = as.Date("2022-07-29"), lty = 2, col = "red")+

  geom_line(col = "gold2", lwd =  0.6)+
  ylab(expression(Surface~light~(W~m^-2)))+
  xlab(NULL) +                                  
  scale_x_date(date_breaks = "1 year",  date_labels = "",  expand = c(0, 0)) +   theme( axis.title.y = element_text(size = 9),
    axis.text.x = element_blank(),  # remove all x-axis labels
    axis.ticks.x = element_line()) +
  scale_alpha_manual(values = c("Before" = 0.6, "After" = 1))+
  theme(legend.position = "null")+
  ylim (c(0,400)) 
 
  
 q <- ggplot(met.mle.final, aes(x = date, y = SV.cms, alpha = fire))+
geom_vline(xintercept = as.Date("2022-08-03"), lty = 2, col = "black")+
  geom_vline(xintercept = as.Date("2022-07-29"), lty = 2, col = "red")+ geom_line(col = "cyan4", lwd =  0.6)+
  ylab(expression(Discharge~(m^3~s^-1)))+
  xlab(NULL) +                                  
  scale_x_date(date_breaks = "1 year",  date_labels = "",  expand = c(0, 0)) +   theme( axis.title.y = element_text(size = 9),
    axis.text.x = element_blank(),  # remove all x-axis labels
    axis.ticks.x = element_line())  +
  scale_alpha_manual(values = c("Before" = 0.6, "After" = 1))+
  theme(legend.position = "null")+
   ylim (c(0,500)) +
   annotate("text",x = as.Date("2022-08-03"),
           y = 420,label = "Before fire &\ndebris flows", hjust = 1.1,color = "red", size = 3, alpha = 0.7, fontface = "bold")+
      annotate("text",x = as.Date("2022-08-03"),
           y = 420,label = "After fire &\ndebris flows", hjust = -0.1,color = "red", size = 3, fontface = "bold")+
annotate("segment", x=as.Date("2022-07-15"), xend=as.Date("2021-09-01"), y=325, yend=325, arrow=arrow(length=unit(0.07,"inches"), ends="last", type="closed"), color="red", linewidth=0.8, alpha = 0.7)+
   annotate("segment", x=as.Date("2022-08-18"), xend=as.Date("2023-05-15"), y=325, yend=325, arrow=arrow(length=unit(0.07,"inches"), ends="last", type="closed"), color="red", linewidth=0.8)
  
 

t <-  ggplot(met.mle.final, aes(x = date, y = turb_median, alpha = fire))+
geom_vline(xintercept = as.Date("2022-08-03"), lty = 2, col = "black")+
  geom_vline(xintercept = as.Date("2022-07-29"), lty = 2, col = "red")+ geom_line(col = "chocolate4", lwd =  0.6)+
  ylab("Turbidity (FNU)")+
  scale_y_log10()+
   xlab(NULL) +                                  
  scale_x_date(date_breaks = "1 year",  date_labels = "",  expand = c(0, 0)) +   theme( axis.title.y = element_text(size = 9),
    axis.text.x = element_blank(), 
    axis.ticks.x = element_line())  +
  scale_alpha_manual(values = c("Before" = 0.6, "After" = 1))+
  theme(legend.position = "null")
 

gpp <-  
  ggplot(met.mle.final, aes(x = date, y = GPP, alpha = fire))+
geom_vline(xintercept = as.Date("2022-08-03"), lty = 2, col = "black")+
  geom_vline(xintercept = as.Date("2022-07-29"), lty = 2, col = "red")+ geom_line(col = "forestgreen", lwd =  0.6)+
  ylab(expression(GPP~(g~O[2]~m^{-2}~d^{-1})))+
   xlab(NULL) +                                  
  scale_x_date(date_breaks = "1 year",  date_labels = "",  expand = c(0, 0)) +   theme( axis.title.y = element_text(size = 9),
    axis.text.x = element_blank(),  
    axis.ticks.x = element_line())+
  scale_alpha_manual(values = c("Before" = 0.6, "After" = 1))+
  theme(legend.position = "null")
 

er <-  ggplot(met.mle.final, aes(x = date, y = ER, alpha = fire))+
geom_vline(xintercept = as.Date("2022-08-03"), lty = 2, col = "black")+
  geom_vline(xintercept = as.Date("2022-07-29"), lty = 2, col = "red")+ geom_line(col = "chocolate", lwd =  0.6)+
  ylab(expression(ER~(g~O[2]~m^{-2}~d^{-1})))+
   xlab(NULL) +                                  
  scale_x_date(date_breaks = "1 year",  date_labels = "",  expand = c(0, 0)) +   theme( axis.title.y = element_text(size = 9),
    axis.text.x = element_blank(),  # remove all x-axis labels
    axis.ticks.x = element_line())  +
  scale_alpha_manual(values = c("Before" = 0.6, "After" = 1))+
  theme(legend.position = "null")
 

nep <-  ggplot(met.mle.final, aes(x = date, y = NEP, alpha = fire))+
geom_vline(xintercept = as.Date("2022-08-03"), lty = 2, col = "black")+
  geom_vline(xintercept = as.Date("2022-07-29"), lty = 2, col = "red")+  geom_hline(yintercept = 0, lty = 1, lwd = 0.2)+
 geom_line(col = "grey20", lwd = 0.6)+
  ylab(expression(NEP~(g~O[2]~m^{-2}~d^{-1})))+
   xlab(NULL) +                                  
  scale_x_date(date_breaks = "1 year",  date_labels = "%Y",  expand = c(0, 0)) +   
  theme( axis.title.y = element_text(size = 9),
    axis.ticks.x = element_line()) +
  scale_alpha_manual(values = c("Before" = 0.6, "After" = 1))+
  theme(legend.position = "null")
 

fire_date <- as.Date("2022-08-03")


ts <- plot_grid(l, q, t, gpp, er, nep, ncol = 1, align = "v", labels = c("a", "b", "c", "d", "e", "f"), label_x = .965, label_y = 0.999, rel_heights = c(1, 1, 1, 1, 1, 1.1))





```


### Calc mean here:

This is just raw data means--not sure why I did this here. I have a table of the means from the model that were calc in the Stats script.

```{r, warning=F, size = 'small', echo=FALSE, fig.width=6.5, fig.height=8}

means_cis <- met.mle.final %>%
  pivot_longer(cols = c(SV.cms, GHI.day, NEP, ER, GPP, turb_median), 
      names_to = "var", values_to = "value") %>%
  group_by(var, fire) %>%
  summarise(mean = mean(value, na.rm = TRUE),
    lower = mean - 1.96 * sd(value, na.rm = TRUE) / sqrt(n()),
    upper = mean + 1.96 * sd(value, na.rm = TRUE) / sqrt(n()),
    .groups = "drop")

means_cis

```

### Right panel, means and laying out the final plot

The data frames for this plot are in the "Stats_D1_Report.qmd" Script. This plot will not run here without running the models in the Stats script first. Need to run these in unison, but the script is long and the model takes a while to run so I didn't want it to clutter this script. 


```{r, warning=F, size = 'small', echo=FALSE, fig.width=6.5, fig.height=8, eval = F}

df.l$group <- factor(df.l$group, levels = c("Before", "After"))

l <- ggplot(df.l, aes(x = group, y = mean, alpha = group)) +
  geom_point(size = 3.5, col = "gold2") +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.25, col = "gold2") +
 # ylim (c(0,400)) +
  scale_alpha_manual(values = c("Before" = 0.6, "After" = 1))+
  theme(legend.position = "null",
      axis.title.y = element_blank(),
      axis.title.x = element_blank(),
      axis.text.x = element_blank())


df.q$group <- factor(df.q$group, levels = c("Before", "After"))

q <- ggplot(df.q, aes(x = group, y = mean, alpha = group)) +
  geom_point(size = 3, col = "cyan4") +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.25, col = "cyan4") +
 #  ylim (c(0,500)) +
  scale_alpha_manual(values = c("Before" = 0.6, "After" = 1))+
  theme(legend.position = "null",
      axis.title.y = element_blank(),
      axis.title.x = element_blank(),
      axis.text.x = element_blank())

df.t$group <- factor(df.t$group, levels = c("Before", "After"))

t <- df.t %>% mutate(lower = ifelse(lower < 0, 0.01, lower))%>%
ggplot( aes(x = group, y = mean, alpha = group)) +
  geom_point(size = 3, col = "chocolate4") +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.25, col = "chocolate4") +
  scale_alpha_manual(values = c("Before" = 0.6, "After" = 1))+
  theme(legend.position = "null",
      axis.title.y = element_blank(),
      axis.title.x = element_blank(),
      axis.text.x = element_blank())+
   scale_y_log10()


df.g$group <- factor(df.g$group, levels = c("Before", "After"))
g <- ggplot(df.g, aes(x = group, y = mean, alpha = group)) +
  geom_point(size = 3, col = "forestgreen") +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.25, col = "forestgreen") +
  #scale_y_log10(limits = c(0.01, 1000))+
  scale_alpha_manual(values = c("Before" = 0.6, "After" = 1))+
  theme(legend.position = "null",
      axis.title.y = element_blank(),
      axis.title.x = element_blank(),
      axis.text.x = element_blank())


df.e$group <- factor(df.e$group, levels = c("Before", "After"))
e <- ggplot( df.e, aes(x = group, y = mean, alpha = group)) +
  geom_point(size = 3, col = "chocolate") +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.25, col = "chocolate") +
  #scale_y_log10(limits = c(0.01, 1000))+
  scale_alpha_manual(values = c("Before" = 0.6, "After" = 1))+
  theme(legend.position = "null",
      axis.title.y = element_blank(),
      axis.title.x = element_blank(),
      axis.text.x = element_blank())

df.n$group <- factor(df.n$group, levels = c("Before", "After"))

n <- ggplot( df.n, aes(x = group, y = mean, alpha = group)) +
    geom_hline(yintercept = 0, lty = 1, lwd = 0.2)+

  geom_point(size = 3, col = "grey10") +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = .25, col = "grey10") +
  #scale_y_log10(limits = c(0.01, 1000))+
  scale_alpha_manual(values = c("Before" = 0.6, "After" = 1))+
  theme(legend.position = "null",
      axis.title.y = element_blank(),
      axis.title.x = element_blank())+
  scale_y_continuous(limits = c(-3.2, 2.4))+
  annotate("text", x = 1, y = 0.2, label = "Autotrophic", vjust = 0, size = 2.5) +
  annotate("text", x = 1.1, y = -0.2, label = "Heterotrophic", vjust = 1, size = 2.5)

ms <- plot_grid(l, q, t, g, e, n, ncol = 1, align = "v", labels = c("g", "h", "i", "j", "k", "l"), label_x = 0.88, label_y = 0.999, rel_heights = c(1, 1, 1, 1, 1, 1.1))


ts.means <- plot_grid(ts, ms, ncol = 2, align = "v", rel_widths = c(0.75, 0.25))

#pdf("Figures/Publication_Plots/F2-ts.pdf", height = 8, width = 6.5)
ts.means
#dev.off()

```

## Figure 3

### Figure 3 panel A: Turbidity by discharge

- Run the model
- Then, run the figure (next code chunk)

Not a variable slope/variable intercept (this is a random effect with groups) like I originally coded it; what we have is an interaction model (outcomes changed little, but good to use the correct model structure). 


$$T = \beta_0 + \beta_1 \times F + \beta_2 \times Q + \beta_3 \times F \times Q + \epsilon$$


```{r, warning=F, size = 'small', echo=FALSE, fig.width=3, fig.height=3, results = 'hide'}

## fit model

intaction_mod <- met.mle.final %>%
  mutate(fire = ifelse(fire == "Before", 0,
                ifelse(fire == "After", 1, NA)))

fit_logturb <- brm(log(turb_median) ~ log(SV.cms) * fire,
  data = intaction_mod,
  family = gaussian(),
  chains = 4, cores = 4, iter = 1000)

summary(fit_logturb)


## work on this data
## code new data using parameters from model above

newdat <- met.mle.final %>%
  mutate(fire = factor(fire, levels = levels(met.mle.final$fire))) %>%
  group_by(fire) %>%
  group_modify(~ tibble(SV.cms = seq(min(.x$SV.cms, na.rm = TRUE),
                                      max(.x$SV.cms, na.rm = TRUE),
                                      length.out = 100))) %>%
  mutate(SV.t.line = ifelse(fire == "Before", -2.60 + log(SV.cms) * 0.96, 
                                             -2.60 + 3.04 + log(SV.cms) * (0.96 -0.51)))%>%
  mutate(SV.t.line = exp(SV.t.line))%>%
  mutate(SV.t.line.upper = ifelse(fire == "Before",  -2.35 + log(SV.cms) * 1.02,
                                  -2.35 + 3.45 + log(SV.cms) * (1.02 -0.41)))%>%
  mutate(SV.t.line.upper = exp(SV.t.line.upper))%>%
  mutate(SV.t.line.lower = ifelse(fire == "Before",  -2.84 + log(SV.cms) *  0.90 ,
                                  -2.84 + 2.64  + log(SV.cms) * ( 0.90  -0.61)))%>%
    mutate(SV.t.line.lower = exp(SV.t.line.lower))



```


```{r, warning=F, size = 'small', echo=FALSE, fig.width=3, fig.height=2.8}
turb.Q <- ggplot() +
    theme_classic()  +
  geom_ribbon(data = newdat, 
                 aes(x = log(SV.cms), ymin = log(SV.t.line.lower), ymax = log(SV.t.line.upper), group = fire),  
                 alpha = 0.3, fill = "grey")+
  geom_point(data = met.mle.final, 
             aes(x = log(SV.cms), y = log(turb_median), color = fire), 
             size = 0.2, alpha = 0.6) +
  ylab("Turbidity (FNU)") +
  xlab(expression(Discharge~(m^3~s^-1))) +
   geom_line(data = newdat, aes(x = log(SV.cms), y = log(SV.t.line), color = fire),
          linewidth = 0.5)+
 scale_color_viridis_d(option = "inferno", end = 0.6, begin = 0.35,
                       name = NULL, labels = c("Before fire", "After fire")) +
 scale_x_continuous(
  breaks = log(c(40, 100, 400)),
  labels = function(x) round(exp(x), 0)) +
scale_y_continuous(
  breaks = log(c(1, 10, 100, 1000)),
  labels = function(y) round(exp(y), 0))+
   theme(#legend.position = c(.78, 0.88), 
     legend.position = "null",
        axis.title.y=element_text(margin=margin(r=-4)))

#pdf("Figures/Publication_Plots/turb-q.pdf", height = 3, width = 3.2)
#turb.Q
#dev.off()
```

### Figure 3 Panel E 
  - This is the GLM segmented (breakpoint) model
  - This figure will not run without the model being run in the Stats script
  - Prediction lines are for Jan and July conditions, based on pre and post fire years
 
#### GLM model for turb to DO

- Additional model specs are in the Models script
- Running this here allows the plot below to render using just this script

```{r, warning=F, size = 'small', echo=FALSE, fig.width=8, fig.height=4}
## data prep
glm.data <- met.mle.final%>% 
  filter(min.DO != -Inf)%>%
  filter(turb_median < 50)%>%
  mutate(min.DO = min.DO*100, t = log(turb_median), 
         q = log(SV.cms), l = log(GHI.day), g = log(GPP))

## model
bp.mod <- glm(min.DO ~ t + q + l,
              data = glm.data,
              na.action = na.omit)

seg_mod <- segmented(bp.mod, seg.Z = ~t, 
                     psi = list("t" = 10))
```





#### Break point plot

```{r, warning=F, size = 'small', echo=FALSE, fig.width=8, fig.height=4}
### Get July and Jan means

jan.turb <- glm.data %>% filter(mon == 1)%>% summarize(range(turb_median)) ## 1.32-36
jul.turb <- glm.data %>% filter(mon == 7)%>% summarize(range(turb_median)) ## 0.85-4.7

l_low  <- glm.data %>% filter(mon == 1) %>% summarize(l = log(mean(GHI.day)))
l_high <- glm.data %>% filter(mon == 7) %>% summarize(l = log(mean(GHI.day)))

q_low  <- glm.data %>% filter(mon == 7) %>% summarize(q = log(mean(SV.cms)))
q_high <- glm.data %>% filter(mon == 1) %>% summarize(q = log(mean(SV.cms)))

## average conditions (not plotted currently)
q_av  <- glm.data  %>% summarize(q = log(mean(SV.cms)))
l_av <- glm.data  %>% summarize(l = log(mean(GHI.day)))

pred_df <- bind_rows(data.frame(
    t = seq(log(1.32), log(36), length.out = 200),
    l = l_low,q = q_high,scenario = "January"),
    data.frame(
    t = seq(log(0.85), log(4.7), length.out = 200),
    l = l_high,q = q_low,scenario = "July"))


## this is code for plotting the average
## data.frame(t = seq(min(glm.data$t), max(glm.data$t), length.out = 200), l = l_av,q = q_av,scenario = "all")
#pred_df$y_hat <- predict(gls_minDO, newdata = pred_df, type = "response")


pred_df$y_hat <- predict(seg_mod, newdata = pred_df, type = "response")

panel.e <- ggplot(pred_df,aes(x=t,y=y_hat))+
geom_hline(yintercept=90,lty=1,col="black")+
annotate("text", y=90, x = 2, label = "Water Quality Threshold", 
              size = 2.8,  col="black", hjust = -0.1, vjust = 1.4)+
annotate("text", y=105, x = 1.25, label = "Turbidity Breakpoint = 3.6 FNU", 
              size = 2.8,  col="black", hjust = 1.0, vjust = 0)+ 
  
  
geom_vline(xintercept=1.28,lty=2)+
geom_point(data=glm.data,aes(x=t,y=min.DO,col=fire),alpha=.6,size=.8,show.legend=FALSE)+
scale_color_viridis_d(option = "inferno", end = 0.6, begin = 0.35,
                       name = NULL) +
new_scale_color()+
geom_line(aes(col=scenario),linewidth=2, alpha = 0.7)+
scale_color_viridis_d(option="mako",begin=.7,end=.3,labels=c("January (low light, high flow)","July (high light, low flow)"),name=NULL)+
scale_x_continuous(breaks=log(c(1,2,4,8,16,32,64)),labels=function(x)round(exp(x),1))+
theme(legend.position=c(.81,.18))+
ylab(expression(Dissolved~O[2]~"("*"% Saturation"*")"))+
xlab("Turbidity (FNU)")
  
 # average July turb pre-fire = 1.7 FNU (0.53 on log scale) min (1.3)
 # average July turb pre-fire = 5.5 FNU (1.7 on log scale) min (0.85)

```


### Figure 3 panes B-D, and final layout

```{r, warning=F, size = 'small', echo=FALSE, fig.width=5.3, fig.height=7.5}

p1 <- met.mle.final %>%
  ggplot(aes(x = GPP, y = ER * -1, col = fire)) +
  annotate("text", y=20, x = 1.28, label = "1:1 Line", 
              size = 3.5,  col="black", hjust = -1.8, vjust = 0)+ 
  theme_classic()+
   geom_abline(intercept = 0, slope = 1, lty = 1) +
  geom_point(size = .3, alpha = 0.6) +
scale_color_viridis_d(option = "inferno", end = 0.6, begin = 0.35,
                       name = NULL, labels = c("Before fire", "After fire")) +
 
labs( x = expression(GPP~(g~O[2]~m^{-2}~d^{-1})),
  y = expression(abs(ER)~(g~O[2]~m^{-2}~d^{-1})))+
  theme(legend.position = "none",
#axis.title = element_text(size = 9),
axis.text = element_text(size = 9),
axis.title.y=element_text(margin=margin(r=-4)))


p2 <- met.mle.final %>%
  ggplot(aes(y = min.DO*100, x = ER * -1, col = fire)) +
  geom_hline(yintercept=90,lty=1,col="black", lwd = 0.4)+
annotate("text", y=90, x = 2, label = " Water Quality\nThreshold", 
              size = 2.7,  col="black", hjust = 0.5, vjust = 1.4)+
    theme_classic()+
  geom_point(size = .3, alpha = 0.6) +
 scale_color_viridis_d(option = "inferno", end = 0.6, begin = 0.35,
                       name = NULL, labels = c("Before fire", "After fire")) +
labs( y=expression(Dissolved~O[2]~"(" * "% Saturation" * ")"),
  x = expression(abs(ER)~(g~O[2]~m^{-2}~d^{-1})))+
  theme(legend.position = "none",
#axis.title = element_text(size = 9),
axis.text = element_text(size = 9),
axis.title.y=element_text(margin=margin(r=-4)))+
  scale_y_continuous(limits = c(75, 101))


p3 <- met.mle.final %>%
  ggplot(aes(y = GPP, x = turb_median , col = fire)) +
  theme_classic()  +
  geom_point(size = .3, alpha = 0.6) +
  scale_color_viridis_d(option = "inferno", end = 0.6, begin = 0.35,
                        name = NULL, labels = c("Before fire", "After fire")) +
 guides(color = guide_legend(override.aes = list(size = 2, alpha = 0.7))) +
theme(legend.position = c(0.8, 0.6),
  legend.key.height = unit(0.3, "cm"),
  legend.spacing.y = unit(0.1, "cm"))+
  labs(y = expression(GPP~(g~O[2]~m^{-2}~d^{-1})),
       x = "Turbidity (FNU)") +
  theme(legend.position = c(0.8, 0.6),
        axis.text = element_text(size = 9),
        axis.title.y = element_text(margin = margin(r = -4))) +
  scale_x_log10() +
  scale_y_continuous(limits = c(0.02, 22))

top <- plot_grid(turb.Q, p3, p1, p2, ncol = 2, align = "hv",
                 labels = c("a", "b", "c", "d"), 
                 label_x = 0.93, label_y = 0.999)


#pdf("Figures/Publication_Plots/panels.pdf", height = 7.5, width = 5.3 )
plot_grid(top, panel.e, nrow = 2, rel_heights = c(.65, .34),  labels = c(" ", "e"), 
                 label_x = 0.965, label_y = 0.999)
#dev.off() 

```




# Suppliment Figures

Figures S1 and S2 were pulled from the depth script and the metabolism output script. 

## Figure S3: Full metab data set

```{r, warning=F, size = 'small', echo=FALSE, fig.width=8, fig.height=6}

si.met <- ggplot(met.mle.final, aes(x = jday, y = GPP))+
  geom_abline(slope = 0, intercept = 0, width = 1, alpha = 0.6)+
  geom_line(col = "cyan4", alpha = 0.8)+
  geom_ribbon(aes(ymin = GPP.lower, ymax = GPP.upper), 
              alpha = 0.4, fill = "cyan4")+
  geom_line(aes(x = jday, y = ER),
            col = "chocolate3", alpha = 0.8)+
  geom_ribbon(aes(ymin = ER.lower, ymax = ER.upper), 
              alpha = 0.4, fill = "chocolate3")+ 
  geom_point(aes(x = jday, y = NEP),
             alpha = 0.5, size = 0.6)+
  facet_wrap(~year, ncol = 2)+
  xlab("Day of year")+ 
  ylab(expression(Metabolism~(g~O[2]~m^{-2}~d^{-1})))

pdf("Figures/Publication_Plots/SI_Metab.pdf", height = 6, width = 8)
plot_grid(si.met)
dev.off()

```


## Figure S4: full metab data set for the ~month around the time of the fire

2 wks before and 2 wks after fire

```{r, warning=F, size = 'small', echo=FALSE, fig.width=5, fig.height=7}

met <- met.mle.final.FIRE%>%
  filter(date > "2022-07-15" & date < "2022-08-15")%>%
  ggplot(aes(x = date, y = GPP))+
  geom_vline(xintercept = as.Date("2022-07-29"), lty = 2, lwd = 0.2, col="red")+
  geom_vline(xintercept = as.Date("2022-08-03"), lty = 2, lwd = 0.2)+
  geom_abline(slope = 0, intercept = 0, width = 1, alpha = 0.6)+
  geom_point(col = "cyan4", alpha = 0.8)+
  geom_errorbar(aes(ymin = GPP.lower, ymax = GPP.upper), 
              alpha = 0.4, fill = "cyan4")+
  geom_point(aes(x = date, y = ER),
            col = "chocolate3", alpha = 0.8)+
  geom_errorbar(aes(ymin = ER.lower, ymax = ER.upper), 
              alpha = 0.4, fill = "chocolate3")+ 
  geom_point(aes(x = date, y = NEP),
             alpha = 0.5, size = 0.6)+
  xlab("")+ 
  ylab(expression(Metabolism~(g~O[2]~m^{-2}~d^{-1})))+
  annotate("text", y=-50, x = as.Date("2022-07-29"), label = "Fire Start\n29 July 2022", 
              size = 4,  col="red", hjust = 1.2, vjust = 0)+
annotate("text", y=-50, x = as.Date("2022-08-04"), label = "Debris Flow\n03 August 2022", 
              size = 4,  col="black", hjust = 0, vjust = 0)+
  scale_y_continuous(breaks = c(20, 0, -20, -40, -60, -80))


tur <- hour.turb.all%>%
  filter(solar.time > as.POSIXct("2022-07-15 18:00:00") & 
           solar.time < as.POSIXct("2022-08-14 12:30:00"))%>%
  ggplot(aes(x = solar.time, y = turb))+
  geom_vline(xintercept = as.POSIXct("2022-07-29 8:15:00"), 
           lty = 2, lwd = 0.2, col = "red") +
geom_vline(xintercept = as.POSIXct("2022-08-03 05:30:00"), 
           lty = 2, lwd = 0.2)+
  geom_line(col = "chocolate4")+
  xlab("")+ 
  ylab("Turbidity (FNU)")+
  scale_y_log10()

do <- sub.day.do%>%
  filter(datetime > as.POSIXct("2022-07-15 18:00:00") & 
           datetime < as.POSIXct("2022-08-14 12:30:00"))%>%
  ggplot(aes(x = datetime, y = DO.obs))+
  geom_vline(xintercept = as.POSIXct("2022-07-29 8:15:00"), 
           lty = 2, lwd = 0.2, col = "red") +
geom_vline(xintercept = as.POSIXct("2022-08-03 05:30:00"), 
           lty = 2, lwd = 0.2)+
  geom_line(col = "skyblue")+
  xlab("")+ 
  ylab("Dissolved oxygen (mg/L)")


pdf("Figures/Publication_Plots/SI_Metab_fire.pdf", height = 5.5, width = 6)
plot_grid(met, tur, do, nrow = 3, rel_heights = c(.4, .3, .3))
dev.off()



```



## Figure S5 GPP x ER


```{r, warning=F, size = 'small', echo=FALSE, fig.width=6, fig.height=6.5}
facet_labels <- met.mle.final %>%
  group_by(year) %>%
  summarize(
    GPP = min(GPP, na.rm = TRUE) + 0.1 * diff(range(GPP, na.rm = TRUE)),
    ER = max(ER * -5, na.rm = TRUE) - 0.1 * diff(range(ER * -5, na.rm = TRUE)))

xmax <- max(met.mle.final$GPP, na.rm = TRUE)

poly_het <- data.frame(x = c(-2.7, xmax, xmax), y = c(-2.7, -2.7, xmax))

label_df <- data.frame(
  year = 2023,
  label = c("Heterotrophic", "Autotrophic"),
  x = c(Inf, -Inf),
  y = c(Inf, -Inf))

p1 <- met.mle.final %>%
  ggplot(aes(x = GPP, y = ER * -1, col = fire)) +
  geom_polygon(data = poly_het,
    aes(x = x, y = y),
    inherit.aes = FALSE,
    fill = "grey", alpha = 0.8) +
  geom_point(size = 1, alpha = 0.6) +
  geom_abline(intercept = 0, slope = 1) +
  facet_wrap(~year) +
  scale_color_viridis_d(option = "mako", begin = 0.7, end = 0.2) +
  theme( legend.position = "none",
    strip.background = element_blank(),
    strip.text = element_blank(),
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.4),
    axis.line = element_line(colour = "black", linewidth = 0.4)) +
 labs(
    x = expression(GPP~(g~O[2]~m^{-2}~d^{-1})),
    y = expression(abs(ER)~(g~O[2]~m^{-2}~d^{-1})))+
geom_text(data = label_df,
          aes(x = x, y = y, label = label),
          inherit.aes = FALSE,
          hjust = c(1.5, -0.8),
          vjust = c(2, -2))


p2 <- met.mle.final %>%
  mutate(fire = ifelse(fire == "Before", "Before fire", "After fire"))%>%
  ggplot(aes(y = GPP, x = jday, col = fire))+
  geom_point(size = 1, alpha = 0.6, aes(group = year))+
  scale_color_viridis_d(option = "mako", begin = 0.2, end = 0.7)+
  labs(y = expression(GPP~(g~O[2]~m^{-2}~d^{-1})),
    y = "Date")+
  stat_smooth(se = F)+
  scale_x_continuous(
    breaks = c(1, 32, 60, 91, 121, 152, 182, 213, 244, 274, 305, 335),
    labels = month.abb)+
    theme(legend.position = c(0.9, 0.74),
          legend.title = element_blank())+
  xlab(" ")


pdf("Figures/Publication_Plots/erXgpp.pdf")
plot_grid(p1, p2, nrow = 2, rel_heights = c(0.65, 0.4))
dev.off() 


```



## Figure S6: DO during turbidity spikes

###  Middle panel: show the Turb spikes post fire

  - Panel showing daily median turbidity following the fire
  - Numbers show spikes over 50; these are the "events"
  - Post processed number placement in pdf editor

```{r, warning=F, size = 'small', echo=FALSE, fig.width=6.5, fig.height=2.2}

spike.fig <- met.mle.final %>%
  filter(fire == "After") %>%
  {ggplot(., aes(y = turb_median+0.1, x = date)) +
     geom_hline(yintercept = 50, lty = 2, lwd = 0.3) +
     geom_area(fill = "grey60", alpha = 0.5) +
     geom_text(data = data.frame(
                 x = seq(min(.$date), max(.$date)-2, length.out = 8),
                 y = rep(200, 8),
                 label = 1:8,
                 col = viridis(8, option = "inferno", end = 0.8, begin = 0.2)),
               aes(x = x, y = y, label = label, color = col),
               size = 5, fontface = "bold", show.legend = FALSE) +
     scale_color_identity() +
     scale_y_log10(expand = c(0,0)) +
     scale_x_date(limits = c(min(.$date), max(.$date)-2),
                  expand = c(0.005,0.005),
                  date_breaks = "2 months",
                  date_labels = "%b %Y") +
     ylab("Daily median turbidity (NTU)") +
     xlab(" ") +
     theme(axis.title.y = element_text(hjust = +.8),
           plot.margin = margin(t=2,r=2,b=-2,l=2)) }

```
### And the DO plots for each "event"

#### Data prep first

  - There are two sets of plots
  - One goes above the turbidity time series, one below
  - Plots were a total hack to get turbidity and DO on same panel
  - I had to select the "windows"--3 days--of high turbidity events
  - Then, select the hourly data from the metab model output
  - And, the turbidity sub-daily data

```{r, warning=F, size = 'small', echo=FALSE, fig.width=6.5, fig.height=3.4}

t <- met.mle.final %>% 
  dplyr::select(date, turb_median, fire)

spike_dates <- mod.fits.bin %>% 
  left_join(t, by = "date") %>%
  filter(fire == "After", turb_median > 50) %>%
  group_by(year, mon) %>%
  summarize(
    date_of_max = date[which.max(turb_median)],
    max_turb = max(turb_median, na.rm = TRUE),
    .groups = "drop") %>%
  filter(date_of_max != as.Date("2022-12-31"))%>%
  mutate(date_of_max = case_when(date_of_max == as.Date("2023-08-22") ~ as.Date("2023-08-21"), TRUE ~ date_of_max))

window_dates <- spike_dates %>%
  mutate(event = row_number()) %>%
  mutate(date_before = date_of_max - 1,
    date_after  = date_of_max + 1) %>%
  pivot_longer(cols = c(date_before, date_of_max, date_after),
    values_to = "date",
    names_to = "window") %>%
  dplyr::select(date, event) %>%
  distinct() %>%
  arrange(date)

spikes <- mod.fits.bin %>%
  mutate(solar.time = round_date(solar.time, unit = "15 minutes")) %>%
  filter(date %in% window_dates$date)%>%
  left_join(window_dates, by = "date")%>%
  left_join(hour.turb, by = "solar.time")%>%
  mutate(cs.turb  = scale(log(turb+1), center = TRUE, scale = TRUE),
         cs.turb.pos = cs.turb - min(cs.turb, na.rm = TRUE))%>%
  mutate(turb_sc_DO = scales::rescale(turb, to = c(min(DO.obs), max(DO.obs))))

```


#### Set of 4 top panels:

```{r, warning=F, size = 'small', echo=FALSE, fig.width=8, fig.height=2}

df <- spikes %>% filter(event == 1)
#df %>% filter(turb == max(turb, na.rm = T)) ## 1521

p_turb <- ggplot(df, aes(x = solar.time)) +
  geom_area(aes(y = turb), fill = "grey80", color = "grey80", alpha = 0.6) +
  theme_void()+
  theme(plot.margin = margin(0.1, 0.0, 0.5, 1.2, unit = "cm"))+
  annotate("text",
           x = df$solar.time[which.max(df$turb)],
           y = 1500,
           label = "1500 NTU", size = 3, color = "grey30", fontface = "bold",
           hjust = -0.15, vjust = 1)

p_DO <- ggplot(df, aes(x = solar.time)) +
  geom_hline(yintercept = 6, lty = 2, col = "red", lwd = 0.5)+
  geom_point(aes(y = DO.obs), alpha = 0.4, size = 0.9, color = "cyan3") +
  geom_line(aes(y = DO.mod), color = "cyan4", size = 0.4) +
 scale_x_datetime(expand = c(0,0))+
  theme_classic() +
  theme(panel.background = element_rect(fill = "transparent", color = NA),
    plot.background = element_rect(fill = "transparent", color = NA),
    #axis.text.x = element_blank(),
    axis.title.x = element_blank())+
scale_y_continuous(name = expression("Dissolved O"[2]*" (mg/L)"), labels = scales::label_number(accuracy = 0.1))

e1 <- ggdraw() +
  draw_plot(p_turb) +
  draw_plot(p_DO, 0, 0, 1, 1)


df <- spikes %>% filter(event == 2)
#df %>% filter(turb == max(turb, na.rm = T)) ## 216

p_turb <- ggplot(df, aes(x = solar.time)) +
  geom_area(aes(y = turb), fill = "grey80", color = "grey80", alpha = 0.6) +
  theme_void()+
  theme(plot.margin = margin(0.1, 0.0, 0.5, 0.6, unit = "cm"))+
  annotate("text",
           x = df$solar.time[which.max(df$turb)],
           y = 220,
           label = "220 NTU", size = 3, color = "grey30", fontface = "bold",
           hjust = -0.15, vjust = 1)

p_DO <- ggplot(df, aes(x = solar.time)) +
  geom_point(aes(y = DO.obs), alpha = 0.4, size = 0.9, color = "cyan3") +
  geom_line(aes(y = DO.mod), color = "cyan4", size = 0.4) +
  scale_x_datetime(expand = c(0,0))+
  theme_classic() +
  theme(panel.background = element_rect(fill = "transparent", color = NA),
    plot.background = element_rect(fill = "transparent", color = NA),
   # axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank())

e2 <- ggdraw() +
  draw_plot(p_turb) +
  draw_plot(p_DO, 0, 0, 1, 1)


df <- spikes %>% filter(event == 3)
#df %>% filter(turb == max(turb, na.rm = T)) ## 1337

p_turb <- ggplot(df, aes(x = solar.time)) +
  geom_area(aes(y = turb), fill = "grey80", color = "grey80", alpha = 0.6) +
  theme_void()+
  theme(plot.margin = margin(0.1, 0.0, 0.5, 0.8, unit = "cm"))+
  annotate("text",
           x = df$solar.time[which.max(df$turb)],
           y = 1300,
           label = "1300 NTU", size = 3, color = "grey30", fontface = "bold",
           hjust = -0.15, vjust = 1)

p_DO <- ggplot(df, aes(x = solar.time)) +
  geom_point(aes(y = DO.obs), alpha = 0.4, size = 0.9, color = "cyan3") +
  geom_line(aes(y = DO.mod), color = "cyan4", size = 0.4) +
  scale_x_datetime(expand = c(0,0))+
  theme_classic() +
  theme(panel.background = element_rect(fill = "transparent", color = NA),
    plot.background = element_rect(fill = "transparent", color = NA),
  #  axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank())+
scale_y_continuous(labels = scales::label_number(accuracy = 0.1))

e3 <- ggdraw() +
  draw_plot(p_turb) +
  draw_plot(p_DO, 0, 0, 1, 1)


df <- spikes %>% filter(event == 4)
#df %>% filter(turb == max(turb, na.rm = T)) ## 143

p_turb <- ggplot(df, aes(x = solar.time)) +
  geom_area(aes(y = turb), fill = "grey80", color = "grey80", alpha = 0.6) +
  theme_void()+
  theme(plot.margin = margin(0.1, 0.08, 0.5, 0.8, unit = "cm"))+
  annotate("text",
           x = df$solar.time[which.max(df$turb)],
           y = 140,
           label = "140 NTU", size = 3, color = "grey30", fontface = "bold",
           hjust = -0.15, vjust = 1)

p_DO <- ggplot(df, aes(x = solar.time)) +
  geom_point(aes(y = DO.obs), alpha = 0.4, size = 0.9, color = "cyan3") +
  geom_line(aes(y = DO.mod), color = "cyan4", size = 0.4) +
  scale_x_datetime(expand = c(0,0))+
  theme_classic() +
  theme(panel.background = element_rect(fill = "transparent", color = NA),
    plot.background = element_rect(fill = "transparent", color = NA),
   # axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank())+
      theme(plot.margin = margin(0.19, 0.27, 0.19, 0.19, unit = "cm"))

e4 <- ggdraw() +
  draw_plot(p_turb) +
  draw_plot(p_DO, 0, 0, 1, 1)

top <- plot_grid(e1, e2, e3, e4, nrow = 1, rel_widths = c(0.25, 0.22, 0.22, 0.225))
```

#### Set of 4 bottom panels:

```{r, warning=F, size = 'small', echo=FALSE, fig.width=8, fig.height=2}

df <- spikes %>% filter(event == 5)
#df %>% filter(turb == max(turb, na.rm = T)) ## 280

p_turb <- ggplot(df, aes(x = solar.time)) +
  geom_area(aes(y = turb), fill = "grey80", color = "grey80", alpha = 0.6) +
  theme_void()+
  theme(plot.margin = margin(0.1, 0.0, 0.5, 1.3, unit = "cm"))+
  annotate("text",
           x = df$solar.time[which.max(df$turb)],
           y = 280,
           label = "280 NTU", size = 3, color = "grey30", fontface = "bold",
           hjust = -0.15, vjust = 1)

p_DO <- ggplot(df, aes(x = solar.time)) +
  geom_point(aes(y = DO.obs), alpha = 0.4, size = 0.9, color = "cyan3") +
  geom_line(aes(y = DO.mod), color = "cyan4", size = 0.4) +
scale_y_continuous(name = expression("Dissolved O"[2]*" (mg/L)")) +  scale_x_datetime(expand = c(0,0))+
  theme_classic() +
  theme(panel.background = element_rect(fill = "transparent", color = NA),
    plot.background = element_rect(fill = "transparent", color = NA),
    #axis.text.x = element_blank(),
    axis.title.x = element_blank())

e5 <- ggdraw() +
  draw_plot(p_turb) +
  draw_plot(p_DO, 0, 0, 1, 1)


df <- spikes %>% filter(event == 6)
#df %>% filter(turb == max(turb, na.rm = T)) ## 2400

p_turb <- ggplot(df, aes(x = solar.time)) +
  geom_area(aes(y = turb), fill = "grey80", color = "grey80", alpha = 0.6) +
  theme_void()+
  theme(plot.margin = margin(0.1, 0.0, 0.5, 0.6, unit = "cm"))+
  annotate("text",
           x = df$solar.time[which.max(df$turb)],
           y = 2400,
           label = "2400 NTU", size = 3, color = "grey30", fontface = "bold",
           hjust = -0.15, vjust = 1)

p_DO <- ggplot(df, aes(x = solar.time)) +
  geom_point(aes(y = DO.obs), alpha = 0.4, size = 0.9, color = "cyan3") +
  geom_line(aes(y = DO.mod), color = "cyan4", size = 0.4) +
  scale_x_datetime(expand = c(0,0))+
  theme_classic() +
  theme(panel.background = element_rect(fill = "transparent", color = NA),
    plot.background = element_rect(fill = "transparent", color = NA),
   # axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank())+
scale_y_continuous(labels = scales::label_number(accuracy = 0.1))

e6 <- ggdraw() +
  draw_plot(p_turb) +
  draw_plot(p_DO, 0, 0, 1, 1)


df <- spikes %>% filter(event == 7)
#df %>% filter(turb == max(turb, na.rm = T)) ## 1900

p_turb <- ggplot(df, aes(x = solar.time)) +
  geom_area(aes(y = turb), fill = "grey80", color = "grey80", alpha = 0.6) +
  theme_void()+
  theme(plot.margin = margin(0.1, 0.0, 0.5, 0.8, unit = "cm"))+
  annotate("text",
           x = df$solar.time[which.max(df$turb)],
           y = 1900,
           label = "1900 NTU", size = 3, color = "grey30", fontface = "bold",
           hjust = -0.15, vjust = 1)

p_DO <- ggplot(df, aes(x = solar.time)) +
    geom_hline(yintercept = 6, lty = 2, col = "red", lwd = 0.5)+
  geom_point(aes(y = DO.obs), alpha = 0.4, size = 0.9, color = "cyan3") +
  geom_line(aes(y = DO.mod), color = "cyan4", size = 0.4) +
  scale_x_datetime(expand = c(0,0))+
  theme_classic() +
  theme(panel.background = element_rect(fill = "transparent", color = NA),
    plot.background = element_rect(fill = "transparent", color = NA),
  #  axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank())+
scale_y_continuous(labels = scales::label_number(accuracy = 0.1))

e7 <- ggdraw() +
  draw_plot(p_turb) +
  draw_plot(p_DO, 0, 0, 1, 1)


df <- spikes %>% filter(event == 8)
#df %>% filter(turb == max(turb, na.rm = T)) ## 780

p_turb <- ggplot(df, aes(x = solar.time)) +
  geom_area(aes(y = turb), fill = "grey80", color = "grey80", alpha = 0.6) +
  theme_void()+
  theme(plot.margin = margin(0.1, 0.08, 0.5, 0.8, unit = "cm"))+
  annotate("text",
           x = df$solar.time[which.max(df$turb)],
           y = 780,
           label = "780 NTU", size = 3, color = "grey30", fontface = "bold",
           hjust = -0.05, vjust = 1)

p_DO <- ggplot(df, aes(x = solar.time)) +
  geom_point(aes(y = DO.obs), alpha = 0.4, size = 0.9, color = "cyan3") +
  geom_line(aes(y = DO.mod), color = "cyan4", size = 0.4) +
  scale_x_datetime(expand = c(0,0))+
  theme_classic() +
  theme(panel.background = element_rect(fill = "transparent", color = NA),
    plot.background = element_rect(fill = "transparent", color = NA),
   # axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank())+
scale_y_continuous(labels = scales::label_number(accuracy = 0.1))+
    theme(plot.margin = margin(0.19, 0.27, 0.19, 0.19, unit = "cm"))

e8 <- ggdraw() +
  draw_plot(p_turb) +
  draw_plot(p_DO, 0, 0, 1, 1)


bottom <- plot_grid(e5, e6, e7, e8, nrow = 1, rel_widths = c(0.25, 0.22, 0.22, 0.225))
```

#### Put all plots together for this overly complex figure

```{r, warning=F, size = 'small', echo=FALSE, fig.width=8, fig.height=6}

#pdf("Figures/Publication_Plots/F5-events.pdf", height = 6, width = 8)
plot_grid(top, spike.fig, bottom, nrow = 3, 
          rel_heights = c(.34, .4, .34))
#dev.off()
```



## Figure S7: GPP by Flow, Turb, and LIght

```{r, warning=F, size = 'small', echo=FALSE, fig.width=6.5, fig.height=7.8}
fire_day <- met.mle.final%>% filter(date == "2022-08-03")
before_day <- met.mle.final%>% filter(date == "2022-08-02")


q <- met.mle.final %>%
  ggplot(aes(x = SV.cms, y = GPP, col = jday, alpha = fire))+
  geom_point(size = 0.5)+
  geom_path(lwd = 0.5) +
  facet_wrap(~year, ncol = 1)+
  scale_color_viridis_c()+
  theme_bw()+
  theme(legend.position = "null",
        strip.text = element_blank(),
    strip.background = element_blank(),
     panel.grid = element_blank())+
  ylab(expression(GPP~(g~O[2]~m^{-2}~d^{-1})))+
  xlab(expression(Discharge~(m^3~s^-1))) +
  scale_x_log10()+
  geom_text(aes(x = Inf, y = Inf, label = year),
            hjust = 1.2, vjust = 1.5, size = 4, col = "grey20") +
    scale_alpha_manual(values = c("Before" = 0.3, "After" = 0.9))+
    geom_point(data = fire_day, aes(x = SV.cms, y = GPP), col = "red", size = 2.3, alpha = 0.5)



t <- met.mle.final %>%
  ggplot(aes(x = turb_median, y = GPP, col = jday, alpha = fire))+
   geom_point(size = 0.5)+
  geom_path(lwd = 0.5) +
  facet_wrap(~year, ncol = 1)+
  scale_color_viridis_c()+
  theme_bw()+
theme(legend.position = "null",
      axis.title.y = element_blank(),
      axis.text.y = element_blank(),
      strip.text = element_blank(),
    strip.background = element_blank(),
     panel.grid = element_blank(),
    axis.title.x = element_text(margin = margin(t = 6)))+
  xlab("Turbidity (FNU)")+
  scale_x_log10()+
    scale_alpha_manual(values = c("Before" = 0.3, "After" = 0.9))+
    geom_point(data = fire_day, aes(x = turb_median, y = GPP), col = "red", size = 2.3, alpha = 0.5)


l <- met.mle.final %>%
  ggplot(aes(x = GHI.day, y = GPP, col = jday, alpha = fire))+
   geom_point(size = 0.5)+
  geom_path(lwd = 0.5) +
  facet_wrap(~year, ncol = 1)+
  scale_color_viridis_c(name = "Day of year")+
  theme_bw()+
  theme(legend.position = c(0.35, 0.108),
       legend.direction = "horizontal",
      axis.title.y = element_blank(),
      axis.text.y = element_blank(),
      strip.text = element_blank(),
    strip.background = element_blank(),
     panel.grid = element_blank())+
  scale_x_continuous(breaks = pretty)+
  xlab(expression(Surface~light~(W~m^-2)))+
  scale_x_log10()+
  guides(colour = guide_colorbar(
    title.position = "top",
    title.hjust = 0.5,
    barwidth = 5,     
    barheight = 0.7   ))+
    scale_alpha_manual(values = c("Before" = 0.3, "After" = 0.9))+
    geom_point(data = fire_day, aes(x = GHI.day, y = GPP), col = "red", size = 2.3, alpha = 0.5)+
  guides(alpha = "none")




#pdf("Figures/Publication_Plots/Drivers.pdf", height = 8, width = 6.5)
plot_grid(q, t, l, ncol = 3, rel_widths = c(.37, .3, .3), align = "h")
#dev.off()

```


## Figure S8: DO min and MEtabolism

```{r, warning=F, size = 'small', echo=FALSE, fig.width=6.5, fig.height=2.5}

p1 <- met.mle.final %>%
  ggplot(aes(x = GPP, y = 100*min.DO, col = fire))+
  geom_hline(yintercept = 90, lty = 2)+
  geom_point(alpha = 0.5, size = 0.5)+
  scale_color_viridis_d(option = "inferno", begin = 0.3, end = 0.6)+
  stat_smooth(method = "lm", se = F)+
  scale_y_continuous(limits = c(75, 105), name = expression("Dissolved O"[2]*" (% Saturation)"))+
  theme(legend.position = "null")+
  xlab(expression(GPP~(g~O[2]~m^{-2}~d^{-1})))

p2 <- met.mle.final %>%
  ggplot(aes(x = ER, y = 100*min.DO, col = fire))+
  geom_hline(yintercept =90, lty = 2)+
  geom_point(alpha = 0.5, size = 0.5)+
  scale_color_viridis_d(option = "inferno", begin = 0.3, end = 0.6,
       labels = c("After" = "After fire","Before" = "Before fire"))+
  stat_smooth(method = "lm", se = F)+
  scale_y_continuous(limits = c(75, 105))+
  theme(legend.position = c(.72, .17), 
        legend.title = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank())+
    xlab(expression(ER~(g~O[2]~m^{-2}~d^{-1})))


p3 <- met.mle.final %>%
  ggplot(aes(x = NEP, y = 100*min.DO, col = fire))+
  geom_vline(xintercept = 0, lty = 2, lwd = .5)+
  geom_hline(yintercept = 90, lty = 2)+
  geom_point(alpha = 0.5, size = 0.5)+
  scale_color_viridis_d(option = "inferno", begin = 0.3, end = 0.6)+
  stat_smooth(method = "lm", se = F)+
  scale_y_continuous(limits = c(75, 105))+
  theme(legend.position = "null",
         axis.title.y = element_blank(),
         axis.text.y = element_blank())+
  xlab(expression(NEP~(g~O[2]~m^{-2}~d^{-1})))

pdf("Figures/Publication_Plots/DO_SAT.pdf", height = 2.5, width = 6.5)
plot_grid(p1, p2, p3, ncol = 3, rel_widths = c(.4, .34, .34))
dev.off() 
```



```{r, warning=F, size = 'small', echo=FALSE, fig.width=6.5, fig.height=2.5}

ig <- readNWISdv("11530500","00060","2010-01-01",Sys.Date()) %>%
  rename(flow = X_00060_00003) %>%
  mutate(date = as.Date(Date),
         jday = yday(date),
         year = year(date),
         dam = ifelse(year < 2025, "dammed", "removed")) %>%
  filter(jday %in% c(160:260))


sv <- readNWISdv("11516530","00060","2010-01-01",Sys.Date()) %>%
  rename(flow = X_00060_00003) %>%
  mutate(date = as.Date(Date),
         jday = yday(date),
         year = year(date),
         dam = ifelse(year < 2025, "dammed", "removed")) %>%
  filter(jday %in% c(60:132))

  
  ggplot(ig, aes(x=jday, y = flow, alpha = dam, group = year))+
  geom_line()+
    scale_y_log10()
  
  
  
  
  
i <- ig %>% group_by(dam)%>%
    summarize (cv = sd(flow) / mean(flow),
               rbi = sum(abs(diff(flow))) / sum(flow))%>%
  mutate(site = "IG")

  
sv %>% group_by(dam)%>%
    summarize (cv = sd(flow) / mean(flow),
               rbi = sum(abs(diff(flow))) / sum(flow))%>%
  mutate(site = "SV")%>%
  rbind(i)

```


